\documentclass[../main.tex]{subfiles}

\begin{document}
\newpage
\renewcommand\abstractname{Acknowledgements}
\begin{abstract}
    \noindent
    
    Having time, curiosity and support which are the requirements to be able to learn something shouldn’t be taken for granted. I had the privilege and circumstances that these factors came together and hope I’m able to appreciate this enough. In the following I want to thank all who made my studies possible. In case I forget to mention someone I want to apologize in advance.  

    First of all, I want to thank my family, parents Manuela and Bernhard and my sister Cornelia, without their unconditional support and background in these turbulent times, studying would not have been possible! 

    Beside the mentioned factors, if you want to learn something you need proper teachers, materials and guidance. And I must say that I really enjoyed the whole program, the classes and exercises and always had the feeling to learn something interesting and useful. Therefore I want to thank all my professors, teachers, tutors and lecturer, especially those from the Institute of Machine Learning. I also want to thank my master thesis supervisors Philipp Seidl and Johannes Schimunek for their guidance and valuable feedback!

    Furthermore, I want to thank MIC, the company where I have worked for over 8 years, and related colleagues and supervisors who initially enabled me to study by providing a flexible and supporting working environment.

    Personally I have the wish that I will be able to use the acquired knowledge appropriately, hopefully benefiting beings, but atleast not harming anyone. Furthermore I hope that I will not forget that behind all data, statistics, methods and algorithms is always a potential affected living being, which has a unique subjective perception different from mine, but similar to myself also strives for happiness and peacefulness!

    



\end{abstract}
\newpage

\end{document}